---
title: "Lab 11 - Grading the professor, Pt. 2"
author: "Insert your name here"
date: "Insert date here"
output: github_document
---

## Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
library(tidymodels)
library(openintro)
library(readxl)
library(titanic)
library(haven)
library(lmtest)
library(DescTools)

titanic3 <- read_excel("data/titanic3.xls",
    col_types = c("numeric", "numeric", "text",
        "text", "numeric", "numeric", "numeric",
        "text", "numeric", "text", "text",
        "text", "text", "text"))


data("titanic_train")
data("titanic_test")
```

## Exercise 1

There are `r nrow(titanic3)` observations and `r ncol(titanic3)` variables in the titanic3 dataset. There is some missing data on age, fare, boat, and body. The data on the primary variables such as class, surival, and sex are complete. 

```{r exercise1_code}
# Rename columns to lowercase
names(titanic3) <- names(titanic3) %>% tolower()
names(titanic_train) <- names(titanic_train) %>% tolower()
names(titanic_test) <- names(titanic_test) %>% tolower()

# Convert pclass to factor

titanic_train <- titanic_train %>%
  mutate(pclass_ord = factor(pclass, ordered = TRUE, levels = c(3, 2, 1)))

titanic_test <- titanic_test %>%
  mutate(pclass_ord = factor(pclass, ordered = TRUE, levels = c(3, 2, 1)))

titanic3 <- titanic3 %>%
  mutate(pclass_ord = factor(pclass, ordered = TRUE, levels = c(3, 2, 1)))

glimpse(titanic3)
summary(titanic3)

titanic3 <- titanic3 %>%
  mutate(sex = case_when(
    sex == "female" ~ 0,
    sex == "male" ~ 1
  ))
```

## Exercise 1.1 
Fit a logistic regression predicting survival from passenger sex and class (as a numeric variable). Save the model as m_apparent

```{r exercise1.1_code}
m_apparent <- glm(survived ~ sex + pclass, data=titanic3, family = "binomial")
summary(m_apparent)
confint(m_apparent)
PseudoR2(m_apparent, which="CoxSnell")
lrtest(m_apparent)
```

## Excercise 1.2 
Now we can use the model to generate predicted survival probabilities for every passenger. For example, p_apparent says that for passenger 1, the likelihood of survival is .89, so out of 100 identical voyages, you'd expect passenger 1 to surive 89 times. 

```{r exercise1.2_code}
p_apparent <- predict(m_apparent, type = "response")
view(p_apparent)
```

## Exercise 1.3 
To turn probabilities into decisions (we’re working for an insurance agency after all), we need a decision rule. To keep things simple, we will use a cutoff of 0.5: if the predicted probability is greater than 0.5, we predict survival (1); otherwise, we predict non-survival (0).

```{r exercise1.3_code}
yhat_apparent <- ifelse(p_apparent > .5, 1, 0)
view(yhat_apparent)
```

##  Exercise 1.4 
Compute the model’s accuracy. We define accuracy as the proportion of correct predictions (both survivors and non-survivors) out of all passengers.

```{r}
acc_apparent <- mean(yhat_apparent == titanic3$survived, na.rm = TRUE)
acc_apparent
```

## Exercise 1.5 Reflection 

 1. Why is apparent accuracy likely to be an overestimate of true predictive performance?
 The accuracy we get when using the same data that created the datafile is probably going to be the best its ever going to get, since the relationship between class, sex, and survival is going to be different if there were a hypothetical titanic type situation to compare it to. Then, the model would be less accurate in its predictions. 
 
 2. Can you think of an analogy from everyday life where “testing on the same data you trained on” would give misleadingly good results?
 
I like to sometimes play the bots on chess.com for practice, but if you play the same bot too many times, you get a sense for the general strategy it assumes, and so you are more likely to beat it and assume you are better than you actually are. 
 
## Exercise 2.1 
Fit the same model as before (survived ~ sex + pclass), but only on the training data. Save it as m_split
 
```{r exercise2.1_code}
m_split <-  glm(survived ~ sex + pclass, data=titanic_train, family = "binomial")
summary(m_split)
```
## Exercise 2.2 
Let’s compute the model’s accuracy on the data it was trained on. This is still “apparent” accuracy, but now limited to the training set.

```{r exercise2.2_code}
p_train <- predict(m_split, type = "response")
yhat_train <- ifelse(p_train > 0.5, 1, 0)

acc_train <- mean(yhat_train == titanic_train$survived, na.rm = TRUE)
acc_train
```

## Exercise 2.3 
Now the real test. We evaluate this model on passengers it has never seen, using titanic_test. Notice the crucial difference: we pass newdata = titanic_test to predict(). This forces the model to make predictions for passengers that played no role in fitting its coefficients.

```{r exercise2.3_code}
p_test <- predict(m_split,
                  newdata = titanic_test,
                  type = "response")

yhat_test <- ifelse(p_test > 0.5, 1, 0)
yhat_test

acc_test <- mean(yhat_test == titanic_test$survived, na.rm = TRUE)
acc_test

```

## Exercise 2.4 Reflection 
1. Which is larger, acc_train or acc_test? Why is that the typical pattern?
acc_train is bigger with 891 obeservations, and the test is on 418 observations. Having more data = better model, so you want to maximize the number of observations used to train the data, while keeping enough data to test on. 

2. Which estimate is closer to what Lloyd’s actually needs — a measure of how well the model explains past data, or how well it predicts future passengers?
We didn't get an accuracy value for the titanic_test dataset, so we don't know. 

3. If acc_test happened to be higher than acc_train, would that invalidate the logic of holdout testing? Explain.
Yes. The point of holdout testing is to have a bigger training dataset so that you have a good model, and then having a smaller thing to test it out on. Like the leave-one-out cross validation that happens in some fmri studies when they are training and testing a classifier. 

## Exercise 3.1 Create Folds
Create titanic_cv by:
  filtering to complete cases on survived, sex, and pclass,
  setting a seed for reproducibility,
  adding a fold variable with values 1 through 10, assigned randomly.
```{r exercise 3.1}
set.seed(100)
titanic_cv <- titanic3 %>%
  filter(!is.na(pclass), !is.na(survived), !is.na(sex)) %>%
  mutate(fold = sample(rep(1:10, length.out = n())))

titanic_cv %>% count(fold)

```

## Exercise 3.2

Now the core of cross validation. Complete the loop so that each fold is used as the test set exactly once. For each iteration:

Split the data: everything except fold j is the training set; fold j is the test set.
Fit the logistic regression on the training set.
Predict on the test set and compute accuracy.
Store the fold accuracies in cv_results.

```{r exercise3.2_code}
cv_results <- data.frame(fold = sort(unique(titanic_cv$fold)), accuracy = NA_real_)

for (j in cv_results$fold) {

  train_j <- titanic_cv %>% filter(fold != j)
  test_j  <- titanic_cv %>% filter(fold == j)

  m_j <- glm(
    formula = survived ~ sex + pclass,
    data = train_j,
    family = binomial
  )

  p_j <- predict(m_j, newdata = test_j, type = "response")
  yhat_j <- ifelse(p_j > 0.5, 1, 0)

  cv_results$accuracy[cv_results$fold == j] <- mean(yhat_j == test_j$survived, na.rm = TRUE)
}

cv_results %>% arrange(fold)

```

## Exercise 3.3
```{r exercise3.3_code}
cv_mean <- mean(cv_results$accuracy)
cv_sd   <- sd(cv_results$accuracy)
cv_min  <- min(cv_results$accuracy)
cv_max  <- max(cv_results$accuracy)

c(cv_mean = cv_mean, cv_sd = cv_sd, cv_min = cv_min, cv_max = cv_max)
```

## Exercise 3.4

```{r exercise3.4_code}
ggplot(cv_results, aes(x = factor(fold), y = accuracy)) +
  geom_col(fill = "steelblue") +
  geom_hline(yintercept = mean(cv_results$accuracy), linetype = "dashed", color = "red") +
  labs(
    title = "10-Fold Cross Validation: Accuracy by Fold",
    x = "Fold",
    y = "Accuracy"
  ) +
  theme_minimal()

```

## Exercise 3.5 Reflection 

1. Why is cv_mean usually lower than acc_apparent? Because the model is more accurate ont he data it was trained with than the data it is tested on. 

2. What does cv_sd tell you that cv_mean does not? The cv_sd tells us how variable the range of prediction accuracy was. 

3. If Lloyd’s demanded a single performance estimate, would you report acc_test (Exercise 2) or cv_mean (Exercise 3)? Defend your choice. I would report cv_mean because it is a composite of 10 testing scenarios, whereas the acc_test is just one value. 

## Exercise 4.1 
```{r exercise4.1_code}
cutoffs <- c(0.3, 0.5, 0.7)

cutoff_results <- data.frame(
  cutoff = cutoffs,
  accuracy = NA_real_
)

for (i in seq_along(cutoffs)) {

  c0 <- cutoffs[i]

  yhat_c <- ifelse(p_test > c0, 1, 0)
  cutoff_results$accuracy[i] <- mean(yhat_c == titanic_test$survived, na.rm = TRUE)

}

cutoff_results

```
